Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2018-10-20 isolation_by_coalescence.tex
another pass through

\begin{abstract}
A number of methods commonly used in landscape genetics
use an analogy to electrical resistance on a network
to describe and fit barriers to movement across the landscape
using genetic distance data.
These are motivated by a mathematical equivalence between electrical resistance 
between two nodes of a network
and the ``commute time'',
which is the mean time for a random walk on that network to leave one node, visit the other, and return.
Here, we describe the differences between commute time
and coalescence time, which more accurately models genetic distances,
and explore the consequences for inference.
We implement a Bayesian method to infer effective movement rates and population sizes 
under both these models,
and find that inference using commute times
can dramatically mislead in the presence of biased gene flow.
We then use forwards-time simulation with continuous geography to demonstrate that
coalescence-based inference remains more accurate than resistance-based methods on realistic data,
but difficulties highlight the need for methods 
that explicitly model continuous, heterogeneous geography.
\end{abstract}

@@

As motivated by \citet{mcrae2006isolation},
commute time (i.e., resistance distance) is a more computationally tractable approximation
to \emph{coalescence time}, a more accurate model of genetic distance data.
Genetic distances between two genomes sampled from the landscape
do derive from an average across a large number of paths between those points --
the lineages along which each segment of genetic material has been inherited
from their most recent common ancestor.
The more recent this ancestor is,
the smaller the genetic distance will be.
Thanks to recombination, there are a large number of such paths, 
and genetic distance averages over these.
It is reasonable in some situations to model these lineages as random walks across the landscape;
however, the model of genetic distance that one is thus led to
is \emph{coalescence time} of the random walks, 
rather than commute time, the quantity that corresponds to resistance.
This naturally raises the question: 
Are methods that depend on effective resistances being misled?
If so, would using coalescence time do better?

\citet{mcrae2006isolation} showed that coalescence time and resistance distance are equal
(up to shift and scaling) in isotropic landscapes such as a ring \citep{matsen2006convergence},
and found that the two were highly correlated in several test cases.
However, all these test cases involved symmetric rates of gene flow,
which is likely not the case in many situations, e.g.,
in rivers \citep{morrissey2009maintenance,sundqvist2016directional,hanks2017modeling},
source-sink dynamics \citep{dias1996sources},
or after population expansions \citep{with2002landscape}.

In this paper, we contrast 
coalescence time and resistance distance,
and develop an inference method for movement rates on a discrete landscape
using both as estimated from genetic distances.
As we will see,
although the two methods are conceptually similar,
commute time inference may get the wrong answer if data are derived from a coalescent process,
especially in the presence of asymmetry.
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information;
however, the geographical resolution is much coarser.
A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem,
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

There are a number of ways to calculate measures of dissimilarity using genetic data.
In this paper we use mean pairwise divergence:
for two genomes, it is calculated as the density of sites at which the two differ.
It is more common in the literature to use $F_{ST}/(1-F_{ST})$,
but the two are highly correlated for good theoretical reasons 
\citet{slatkin1991inbreeding,rousset1997genetic},
especially for SNP data.
The two genomes differ at a site only if there has been a mutation in some ancestor on the lineages leading
from the two genomes back to their most recent common ancestor at that site;
under an infinite-sites model with average mutation rate $\mu$ per site and per meiosis,
pairwise divergence divided by $\mu$
is therefore an unbiased estimate of the mean time to most recent common ancestor,
averaged across the genome \citep{hudson2007variance,ralph2015empirical}.
It is natural to model this process by following the two lineages back through time
(and across space)
until they find their most recent common ancestor,
i.e., until they coalesce.
Suppose we model the geographic distribution of the species
as a collection of randomly mating populations
that exchange occasional migrants,
thus discretizing geographic space.
If the genome is selectively neutral,
then the motion of these lineages back through time
forms a Markov process, the \emph{structured coalescent} \citep{wakeley2005coalescent}.
In this framework, each lineage performs a random walk across the populations,
with movement probabilities depending on the flux of individuals between populations.
This provides the link between genetic distance and random walks.

For data, we are given genotypes of individuals and their geographic coordinates;
then, we divide these individuals into groups,
and compute the genetic distance between group $i$ and group $j$
as the mean divergence between a pair of genomes, chosen one from each group.
We denote the resulting genetic distance $D_{ij}$;
since divergence is symmetric, $D_{ij} = D_{ji}$.
We compute $D_{ii}$ by sampling without replacement from group $i$,
so that generally $D_{ii} > 0$: the diagonal elements measure local diversity.
As discussed below, we will find commute times so that the mean commute time between
$i$ and $j$ matches $D_{ij}$ -- but the commute time of a location to itself is zero,
so we must account for local diversity.
We do this in the same way as \citet{petkova2016visualizing},
by introducing another set of model parameters, $q$, 
that fill an analogous role to the coalescence rates, $\gamma$.
Then, we use commute time to approximate genetic distance as
\begin{align} \label{eq:commute_approx}
	D_{ij} \approx \comdist_{ij} = R_{ij}/4 + q_{i}/2 + q_{j}/2 .
\end{align}
The factor of four appears to make commute time more similar to coalescence time:
the commute time covers the path from $i$ to $j$ twice
(once in both directions),
while the coalescence time covers the distance only once,
and in half the time, because two particles are moving.
Furthermore, coalescence time includes within-location diversity
because particles do not coalesce instantly --
therefore, commute time is most analogous to the first meeting time.
For a conceptual picture of the differences between the coalescence time and commute time,
see Figure \ref{fig:concept_coalcom}.

@@

Although commute times are uniquely determined by movement rates, the reverse is not true.
Commute times only depend on the symmetric part of $H$ --
given a skew-symmetric matrix $Z$ (so that $Z + Z^T = 0$),
any Markov chain that has hitting times given by $H_\epsilon = H + \epsilon Z$ for some $\epsilon$
will have the same commute times.
The resulting hitting times may not be valid --
if we let $G_\epsilon$ be the matrix constructed by applying equation \eqref{eqn:G_from_H} to $H_\epsilon$,
% (assuming this is invertible),
then it is not guaranteed that the offdiagonal elements of $G_\epsilon$ are all nonnegative, as required.
(It is always true that rows of $G_\epsilon$ sum to zero.)
We provide a small concrete example in Appendix \ref{ss:simple_example}.
However, since $G$ and $H$ are continuous functions of each other,
if all offdiagonal elements of $G$ are strictly positive,
then there exists a positive $\epsilon_0$ such that all $G_\epsilon$ for $\epsilon < \epsilon_0$
do define valid Markov chains, all with the same commute times.

@@

This suggests that although nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations.
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
For more discussion of related problems, see 
\citet{epstein2008badtruth,myers2008learn,terhorst2015fundamental}.

@@

Mean coalescence times estimated from real data are subject to a number of sources of noise
(which we discuss in more detail later).
Here, we use
the exact solutions based on linear algebra developed above
to develop a Bayesian inference method that accounts for noise in the data
and estimates uncertainty in the resulting estimates.
To do this, we model genetic distances as equal to coalescence times plus noise:
$D_{ij} = C_{ij} + \epsilon_{ij}$,
where $\epsilon_{ij}$ are independent and Normally distributed
with mean $\mathbf{0}$ and variance $\sigma_\epsilon^2$,
except that $\epsilon_{ij} = \epsilon_{ji}$.
(We imagine that $C_{ij}$ is large compared to $\sigma_{\epsilon}$,
so that resulting values are still positive.)
Using the calculations above, 
given $G$ and $\gamma$ we can compute the corresponding coalescence times
-- which we call $\mathcal{C}(G, \gamma)$ --
so our model is that $D$ has a Gaussian distribution 
with mean $\mathcal{C}(G, \gamma)$.
The maximum likelihood estimate of $G$ could be found directly;
however, this in practice will quite likely contain negative movement rates,
and does not allow us to impose constraints
(such as only allowing a subset of movement rates to be nonzero).

@@

Inference with resistance distance is done in exactly the same way,
except that $q$ replaces $\gamma$,
and instead of $\mathcal{C}(G,\gamma)$ we use $\comdistfn(G, q)$,
which is the matrix $\comdist$ computed from $G$ and $q$ by solving equation \eqref{eq:commute_approx}.

@@

\paragraph{MCMC methods}
To sample from the posterior distribution of equation \eqref{eq:post}, 
we used R \citep{Rmanual} 
to implement a standard Metropolis-Hastings algorithm with Gaussian proposals \citep{brooks2011handbook}.
Starting locations were chosen by sampling from the prior distribution.
Before the usual ``burn-in'' phase before sampling for inference begins,
we included a period of ``pre-burn-in'' which used the same MCMC procedure
with a larger $\sigma_\epsilon$, to allow the chain to more quickly converge on the high-posterior
portion of parameter space.
Typically, we ran MCMC for $10^6$ iterations of pre-burn-in 
and $3 \times 10^6$ iterations of burn-in,
followed by $4 \times 10^6$ iterations which were used to estimate posterior distributions,
although for small graphs fewer iterations were necessary.
We used standard methods to assess convergence and mixing.

@@

We compared inference under coalescence and commute time models
using both data ``from the model'' and more realistic simulations.
For the first category, we designed a number of graphs to test particular aspects of inference,
with migration rates specified on each directed edge
(plotted below using igraph \citep{igraph}).
Given the edge weights ($G$) and the coalescence rates ($\gamma$) of a graph,
we produced data by calculating exact coalescence times ($\mathcal{C}(G,\gamma)$),
to whose entries we added independent Gaussian noise while preserving symmetry,
as described above.
Particular parameters and levels of noise are given in the Results.

These simulations are expected to be nearly equivalent to neutral, discrete population-based simulations
(either forwards-time or coalescent);
the only difference being that noise terms about the theoretical mean should be somewhat correlated.

\paragraph{Forwards-time simulations}
To produce realistic data, we implemented forwards-time simulations using SLiM v3.1 
\citep{haller2018forward,haller2018treesequence},
with individuals living across continuous, two-dimensional geography (sometimes with barriers),
from which we recorded genomic data.
The basic simulation, which we modified to produce several other situations,
is as follows.
Each simulation had around $10^4$ diploid, hermaphroditic individuals, 
living across two-dimensional geography.
Mates were selected from nearby individuals using a Gaussian kernel with standard deviation $\sigma_d$,
truncated at $3 \sigma_d$.
Simulations were done on either a square region with width equal to $80\sigma_d$,
or a $5:3$ rectangular region with the same area.
Offspring are dispersed away from the mother using the same kernel,
reflected from the boundaries of the habitable area.
In each simulation, $\sigma_d$ was chosen 
so that there were an average of 0.25 individuals per $\sigma_d^2$ units of area.
In order to prevent excessive population clumpiness \citep{felsensten1975pain}, 
we also model local competition. 
To do this, we use the same truncated Gaussian kernel to define an ``interaction strength''
between each pair of nearby individuals;
then, the fitness of each individual is set to $\exp(-x/40)$, 
where $x$ is the sum of their interaction strengths with other individuals.
This reduces fitness for individuals in denser areas.
Individuals are chosen to reproduce proportional to their fitness.
Clumpiness is similar to what is seen in 
Figure \ref{fig:5x3b_continuous}.

Individuals have 10 pairs of ``chromosomes'' with $2 \times 10^5$ loci each 
and a recombination rate of $10^{-8}$ per generation per locus.
Neutral mutations were added at a rate of $10^{-9}$ per meiosis.
To make these simulations computationally feasible,
the mutation rate was set to 0.0 during the forwards simulation,
while tree sequences were recorded \citep{haller2018treesequence},
and mutations were added after the fact with msprime,
which is equivalent to including mutations during the forwards simulation 
but resulted in much faster run times \citep{kelleher2018efficient}.

To compute genetic distances,
geographic space was partitioned into square regions,
and within each region, 
individuals are sampled uniformly for ``genotyping'' 
from among those within the middle three quarters of each dimension of the square,
as shown for one situation in Figure \ref{fig:5x3b_continuous}.
This protocol was chosen as a compromise between sampling all individuals 
from near the center of each location,
which could result in a large number of close relatives being sampled, 
leading to an underestimate of the mean genetic distance for individuals in that location,
and uniform sampling from the whole area of each grid, 
which may result in many sampled individuals being very close to other grid squares.
Mean genetic distances for each pair of regions are then computed as described above.
To estimate uncertainty, we computed the standard error of these mean distances
using the minimum number of individuals in each location.

@@

We first tested the method using data generated under the model.
To test the methods across a wide range of situations,
we generated fifty sets of random movement rates on $3 \times 2$ rectangular grids:
twenty-five symmetric (i.e., with $G_{ij} = G_{ji}$) and twenty-five asymmetric.
We generated movement parameters
by rounding independent exponential random variables up to the nearest one tenth,
and fixed coalescence rate at 1 for all locations.
We then added independent noise to each matrix entry (preserving symmetry),
and assume that we can estimate the amount of noise reasonably well, 
so used the standard deviation of the noise for $\sigma_\epsilon$ in the likelihood function.
To assess the impact on inference of the number of parameters,
we performed inference both with and without the assumption 
that coalescence rate is the same everywhere.
For $3 \times 2$ graphs, there are 21 equations
and 14 movement parameters,
so that with a global coalescence rate there are 15 unknowns
but with individual coalescence rates there are 20 unknowns,
which is close to nonidentifiable.
To quantify accuracy across these random graphs in different situations,
we computed, for each model fit on a particular random graph,
the absolute difference between the posterior median of each movement parameter
and the true value.
We then averaged these across movement parameters to produce a measure of accuracy for that model fit,
reported below as ``mean absolute error''.

\paragraph{Varying noise}
Unsurprisingly, coalescence time inference became less accurate as the amount of noise increased.
As we varied the standard deviation of the noise
from 1/1000 to 1/50 of the mean value of $C$ for that grid,
the mean absolute error 
for coalescence time inference with a single coalescence rate
increased from around 0.05 to 0.4.
Posterior interquartile ranges increased proportionally.
(although there was substantial variation in accuracy between replicates;
see Figures \ref{fig:mult_noise} and \ref{fig:mult_noise_iqr}).

Increasing the number of parameters by allowing multiple coalescence rates
made the inference problem even harder --
mean absolute errors no longer depended strongly on the amount of noise added,
and hovered around 0.4.

More surprisingly, the performance of commute time inference did not depend on the amount of noise --
mean absolute error was around 0.6 across all levels of noise,
both with and without more than one parameter for local diversity
(see Figure \ref{fig:mult_noise}).
Although symmetry did not affect coalescence time inference,
commute time inference did substantially worse on asymmetric graphs,
showing median absolute errors of around 0.8.

\paragraph{Varying coalescence rates}
Lower coalescence rates mean that lineages wander about the landscape for longer
before they coalesce, 
thus greatly reducing the geographic information we get from coalescence times.
Indeed, coalescence rates had a strong effect on feasability of the inference problem
(shown in Figures \ref{fig:mult_gam} and \ref{fig:mult_gam_iqr}):
lowering the coalescence rate from 10 to 0.1
(always adding noise with standard deviation 1/200th of the mean value)
increased mean absolute error of coalescence time inference from around 0.05 to 0.5.
Again, allowing multiple coalescence rates
or using commute time inference were highly inaccurate in all cases,
with mean absolute errors of above 0.5.

@@

We then test the method's ability to locate barriers to migration
on the $5 \times 3$ grid with asymmetric gene flow shown in Figure \ref{fig:5x3b_grid}. 
Each nonzero movement rate is determined randomly as before,
and coalescence rates are set to 1 everywhere.
Figure \ref{fig:5x3b_post_coalvcom} shows posterior distributions of the movement rates
for both coalescence time and resistance distance inference
on this graph, with noise standard deviation equal to
1/1000 of the mean coalescence time (left)
and 1/100 of the mean coalescence time (right).
Although increased noise in the data increased uncertainty in the estimated values,
coalescence time inference correctly inferred not only the location of the barrier
but also the migration rates along all other edges within a reasonable margin of error.
Commute time inference also identified the barrier,
but was much less accurate for the other values.
We also explored making the problem harder,
by first (a) dropping data corresponding to populations 2, 5, 7, and 15;
and then (b) allowing a separate coalescence rate for each location,
in both cases using a noise level of 1/100.
The resulting posterior distributions are shown in Figure \ref{fig:5x3boxplots_mult_gam},
and show substantially increased error,
but only in movement rates to the removed locations.
As before, uncertainty was increased when multiple coalescence rates were allowed,
but not as strongly as before, likely because the problem is less close to being underconstrained,
having 105 equations and 59 unknowns.

>>>>

\begin{abstract}
A number of methods commonly used in landscape genetics
use an analogy to electrical resistance on a network
to describe and fit barriers to movement across the landscape
using genetic distance data.
These are motivated by a mathematical equivalence between electrical resistance 
between two nodes of a network
and the ``commute time'',
which is the mean time for a random walk on that network to leave one node, visit the other, and return.
However, genetic data are more accurately modeled by a different quantity,
the coalescence time.
Here, we describe the differences between resistance distance and coalescence time,
and explore the consequences for inference.
We implement a Bayesian method to infer effective movement rates and population sizes 
under both these models,
and find that inference using commute times
can dramatically mislead in the presence of biased gene flow.
We then use forwards-time simulation with continuous geography to demonstrate that
coalescence-based inference remains more accurate than resistance-based methods on realistic data,
but difficulties highlight the need for methods 
that explicitly model continuous, heterogeneous geography.
\end{abstract}

@@

As motivated by \citet{mcrae2006isolation},
commute time (i.e., resistance distance) is a computationally tractable approximation
to the more accurate random walk-based model of genetic distance,
the \emph{coalescence time}.
Genetic distances between two genomes sampled from the landscape
do derive from an average across a large number of paths between those points --
the lineages along which each segment of genetic material has been inherited
from their most recent common ancestor.
The more recent this ancestor is,
the smaller the genetic distance will be.
Thanks to recombination, there are a large number of such paths, 
and genetic distance averages over these.
It is reasonable in some situations to model these lineages as random walks across the landscape;
however, the model of genetic distance that one is thus led to
is \emph{coalescence time} of the random walks, 
rather than commute time, the quantity that corresponds to resistance.
This naturally raises the question: 
Are methods that depend on effective resistances being misled by model misspecification?
If so, would using coalescence time do better?

\citet{mcrae2006isolation} showed that coalescence time and resistance distance are equal
(up to shift and scaling) in isotropic landscapes such as a ring \citep{matsen2006convergence},
and found that the two were highly correlated in several test cases.
However, all test cases had symmetric migration
(the same rate of gene flow in both directions along each edge),
which is likely not the case in many situations, e.g.,
in rivers \citep{morrissey2009maintenance,sundqvist2016directional,hanks2017modeling},
source-sink dynamics \citep{dias1996sources},
or after population expansions \citep{with2002landscape}.

In this paper, we contrast 
coalescence time and resistance distance,
and develop an inference method for movement rates on a discrete landscape
using both as estimated from genetic distances.
As we will see,
although the two methods are conceptually similar,
resistance-based inference may get the wrong answer if data are derived from a coalescent process,
especially in the presence of gene flow asymmetry.
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information;
however, the geographical resolution is much coarser.
(A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.)
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem,
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

There are a number of ways to calculate measures of dissimilarity using genetic data.
In this paper we use mean pairwise divergence:
for two genomes, it is calculated as the density of sites at which the two differ.
It is more common in the literature to use $F_{ST}/(1-F_{ST})$,
but the two are highly correlated for good theoretical reasons 
\citep{slatkin1991inbreeding,rousset1997genetic},
especially for SNP data.
The two genomes differ at a site only if there has been a mutation in some ancestor on the lineages leading
from the two genomes back to their most recent common ancestor at that site.
Therefore,
under an infinite-sites model with average mutation rate $\mu$ per site and per meiosis,
pairwise divergence divided by $\mu$
is an unbiased estimate of the mean time to most recent common ancestor,
averaged across the genome \citep{hudson2007variance,ralph2015empirical}.
It is natural to model this process by following the two lineages back through time
(and across space)
until they find their most recent common ancestor,
i.e., until they coalesce.
Suppose we model the geographic distribution of the species
as a collection of randomly mating populations
that exchange occasional migrants,
thus discretizing geographic space.
If the genome is selectively neutral,
then the motion of these lineages back through time
forms a Markov process, the \emph{structured coalescent} \citep{wakeley2005coalescent}.
In this framework, each lineage performs a random walk across the populations,
with movement probabilities depending on the flux of individuals between populations.
This provides the link between genetic distance and random walks.

For data, we are given genotypes of individuals and their geographic coordinates;
then, we divide these individuals into groups,
and compute the genetic distance between group $i$ and group $j$
as the mean divergence between a pair of genomes, chosen one from each group.
We denote the resulting genetic distance $D_{ij}$;
since divergence is symmetric, $D_{ij} = D_{ji}$.
We compute $D_{ii}$ by sampling without replacement from group $i$,
so that generally $D_{ii} > 0$: the diagonal elements measure local diversity.
As discussed below, we will find commute times so that the mean commute time between
$i$ and $j$ matches $D_{ij}$ -- but the commute time of a location to itself is zero,
so we must account for local diversity.
We do this in the same way as \citet{petkova2016visualizing},
by introducing the \emph{local diversity} parameters $q_i$,
and using the model
\begin{align} \label{eq:commute_approx}
	D_{ij} \approx \comdist_{ij} = R_{ij}/4 + q_{i}/2 + q_{j}/2 .
\end{align}
The factor of four is necessary to make commute time work on the same time scale as coalescence time:
the commute time covers the path from $i$ to $j$ twice
(once in both directions),
while the coalescence time covers the distance only once,
and in half the time, because two particles are moving.
% Furthermore, coalescence time includes within-location diversity
% because particles do not coalesce instantly --
% therefore, commute time is most analogous to the first meeting time.
For a conceptual picture of the differences between the coalescence time and commute time,
see Figure \ref{fig:concept_coalcom}.

@@

Although commute times are uniquely determined by movement rates, the reverse is not true.
Commute times only depend on the symmetric part of $H$ --
given a skew-symmetric matrix $Z$ (so that $Z + Z^T = 0$),
any Markov chain that has hitting times given by $H_\epsilon = H + \epsilon Z$ for some $\epsilon$
will have the same commute times.
The resulting hitting times may not be valid --
if we let $G_\epsilon$ be the matrix constructed by applying equation \eqref{eqn:G_from_H} to $H_\epsilon$,
% (assuming this is invertible),
then it is not guaranteed that the offdiagonal elements of $G_\epsilon$ are all nonnegative, as required.
% (It is always true that rows of $G_\epsilon$ sum to zero.)
We provide a small concrete example in Appendix \ref{ss:simple_example}.
However, since $G$ and $H$ are continuous functions of each other,
if all offdiagonal elements of $G$ are strictly positive,
then there exists a positive $\epsilon_0$ such that all $G_\epsilon$ for $\epsilon < \epsilon_0$
do define valid Markov chains, all with the same commute times.

@@

This suggests that nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations
and direct, long-distance migration is disallowed.
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
For more discussion of related problems, see 
\citet{epstein2008badtruth}, \citet{myers2008learn}, or \citet{terhorst2015fundamental}.

@@

Mean coalescence times estimated from real data are subject to a number of sources of noise
(which we discuss in more detail later).
Here, we use
exact solutions of the equations above
to develop a Bayesian inference method that accounts for noise in the data
and estimates uncertainty in the resulting estimates.
To do this, we model genetic distances as equal to coalescence times plus noise:
$D_{ij} = C_{ij} + \epsilon_{ij}$,
where $\epsilon_{ij}$ are independent and Normally distributed
with mean $\mathbf{0}$ and variance $\sigma_\epsilon^2$,
except that $\epsilon_{ij} = \epsilon_{ji}$.
(We assume that $C_{ij}$ is large compared to $\sigma_{\epsilon}$,
so that resulting values are still positive.)
Given $G$ and $\gamma$ we can solve equation \eqref{eqn:C_matrix} 
to find the corresponding coalescence times
-- which we call $\mathcal{C}(G, \gamma)$ --
so our model is that $D$ has a Gaussian distribution 
with mean $\mathcal{C}(G, \gamma)$.
The maximum likelihood estimate of $G$ could be found directly;
however, this in practice will quite likely contain negative movement rates,
and does not allow us to impose constraints
(such as only allowing a subset of movement rates to be nonzero).

@@

Inference with resistance distance is done in exactly the same way,
except that $q$ replaces $\gamma$,
and instead of $\mathcal{C}(G,\gamma)$ we use $\comdistfn(G, q)$,
which is the matrix $\comdist$ computed from $G$ and $q$ 
by putting the solution to equation \eqref{eqn:hitting_sum}
through equations \eqref{eqn:R_from_H} and \eqref{eq:commute_approx}.

@@

\paragraph{MCMC methods}
To sample from the posterior distribution of equation \eqref{eq:post}, 
we used R \citep{Rmanual} 
to implement a standard Metropolis-Hastings algorithm with Gaussian proposals \citep{brooks2011handbook}.
Starting locations were chosen by sampling from the prior distribution.
Before the usual ``burn-in'' phase,
we included a period of ``pre-burn-in'' which used the same MCMC procedure
with a larger $\sigma_\epsilon$, to allow the chain to more quickly converge on the high-posterior
portion of parameter space.
Typically, we ran MCMC for $10^6$ iterations of pre-burn-in 
and $3 \times 10^6$ iterations of burn-in,
followed by $4 \times 10^6$ iterations which were used to estimate posterior distributions,
although for small graphs fewer iterations were necessary.
We used standard methods to assess convergence and mixing.

@@

We compared inference under coalescence time and resistance distance models
using both data ``from the model'' and from more realistic simulations.
For the first category, we designed a number of graphs to test particular aspects of inference,
with migration rates specified on each directed edge
(plotted below using igraph \citep{igraph}).
Given the edge weights ($G$) and the coalescence rates ($\gamma$) of a graph,
we produced data by calculating exact coalescence times ($\mathcal{C}(G,\gamma)$),
to whose entries we added independent Gaussian noise while preserving symmetry.
We scaled the noise to the mean value of the times themselves,
so that a ``noise level'' of $\alpha$ refers to data in which the standard deviation of the noise
was set to $\alpha$ multiplied by the average coalescence time across all pairs of populations.
Particular parameters and levels of noise are given in the Results.
These simulations are expected to be nearly equivalent to neutral, discrete population-based simulations
(either forwards-time or coalescent),
except that noise terms about the theoretical mean should be somewhat correlated.

\paragraph{Forwards-time simulations}
To produce realistic data, we implemented forwards-time simulations using SLiM v3.1 
\citep{haller2018forward,haller2018treesequence},
with individuals living across continuous, two-dimensional geography (sometimes with barriers),
from which we recorded genomic data.
The basic simulation, which we modified to produce several other situations,
is as follows.
Each simulation had around 6400 diploid, hermaphroditic individuals, 
living across two-dimensional geography.
Mates were selected from nearby individuals using a Gaussian kernel with standard deviation $\sigma_d$,
truncated at $3 \sigma_d$.
Simulations were done on either a square region with width equal to $80\sigma_d$,
or a $5:3$ rectangular region with the same area.
Offspring are dispersed away from the mother using the same kernel,
reflected from the boundaries of the habitable area.
To produce a realistic model and prevent excessive population clumpiness \citep{felsensten1975pain}, 
we model local competition instead of a globally constant population size.
To do this, 
we use the same truncated Gaussian kernel to define an ``interaction strength''
between each pair of nearby individuals, 
and compute the probability of survival of each individual to the next time step as
$\min(0.9, 2 / (1 +  5x/12 \pi))$,
where $x$ is the sum of the interaction strengths of that individual with all other individuals
within distance $3 \sigma_d$.
(The value $x$ is an estimate of local population density.)
This reduces fitness for individuals in denser areas,
and produces an equilibrium population density
of close to one individual per $\sigma_d^2$ units of area.
Clumpiness of the resulting simulations is similar to what is seen in 
Figure \ref{fig:5x3b_continuous}.

Each diploid individual has one pair of homologous chromosomes with $10^8$ loci each 
and a recombination rate of $10^{-8}$ per generation per locus.
Neutral mutations were added at a rate of $10^{-9}$ per meiosis.
To make these simulations computationally feasible,
the mutation rate was set to 0.0 during the forwards simulation,
while tree sequences were recorded \citep{haller2018treesequence},
and mutations were added after the fact with msprime,
which is equivalent to including mutations during the forwards simulation 
but resulted in much faster run times \citep{kelleher2018efficient}.

To compute genetic distances,
geographic space was partitioned into square regions,
and within each region, 
individuals are sampled uniformly for ``genotyping'' 
from among those within the middle three quarters of each dimension of the square,
as shown for one situation in Figure \ref{fig:5x3b_continuous}.
This protocol was chosen as a compromise between sampling all individuals 
from near the center of each location,
which could result in a large number of close relatives being sampled,
and uniform sampling from the whole area of each grid, 
which may result in many sampled individuals being very close to other grid squares.
Mean genetic distances for each pair of regions are then computed as described above.
To estimate uncertainty, we computed the standard error of these mean distances
using the minimum number of individuals in each location.

@@

We first tested the method using data generated under the model.
To test the methods across a wide range of situations,
we generated fifty sets of random movement rates on $3 \times 2$ rectangular grids:
twenty-five symmetric (i.e., with $G_{ij} = G_{ji}$) and twenty-five asymmetric.
We generated movement parameters
by rounding independent exponential random variables up to the nearest one tenth,
and fixed coalescence rates at 1 for all locations.
We then analytically computed mean coalescence times,
and added independent noise to each pairwise coalescence time.
We assume that we can estimate the amount of noise reasonably well, 
so used the true standard deviation of the noise for $\sigma_\epsilon$ in the likelihood function.
To assess the impact on inference of the number of parameters,
we performed inference both with and without the assumption 
that coalescence rate is the same everywhere.
For $3 \times 2$ graphs, there are 21 equations
and 14 movement parameters,
so that with a global coalescence rate there are 15 unknowns
but with individual coalescence rates there are 20 unknowns,
which is close to nonidentifiable.
To quantify accuracy across these random graphs in different situations,
we computed, for each model fit on a particular random graph,
the absolute difference between the posterior median of each movement parameter
and the true value.
We then averaged these across movement parameters to produce a measure of accuracy for that model fit,
reported below as ``mean absolute error''.

\paragraph{Varying noise}
Unsurprisingly, coalescence time inference became less accurate as the amount of noise increased.
As we varied the standard deviation of the noise
from 1/1000 to 1/50 of the mean value of $C$ for that grid,
the mean absolute error of inferred gene flow rates ($G_{ij}$) 
for coalescence time inference with a single coalescence rate
increased from around 0.05 to 0.4.
Since true values of $G_{ij}$ were of order 1,
this is a transition from precise to very rough (but still informative) inference.
Posterior interquartile ranges increased proportionally,
although there was substantial variation in accuracy between replicates;
see Figures \ref{fig:mult_noise} and \ref{fig:mult_noise_iqr}.
Increasing the number of parameters by allowing multiple coalescence rates
made the inference problem even harder --
mean absolute errors no longer depended strongly on the amount of noise added,
and hovered around 0.4.

More surprisingly, the performance of resistance distance inference 
did not depend on the amount of noise --
mean absolute error was around 0.6 across all levels of noise,
both with and without more than one parameter for local diversity
(see Figure \ref{fig:mult_noise}).
Although symmetry did not affect coalescence time inference,
resistance distance inference did substantially worse on asymmetric graphs,
showing median absolute errors of around 0.8 --
inferred gene flow rates were almost uncorrelated with true values.

\paragraph{Varying coalescence rates}
Lower coalescence rates mean that lineages wander about the landscape for longer
before they coalesce, 
thus greatly reducing the geographic information we get from coalescence times.
Indeed, coalescence rates had a strong effect on feasibility of the inference problem
(shown in Figures \ref{fig:mult_gam} and \ref{fig:mult_gam_iqr}):
lowering the coalescence rate from 10 to 0.1
(always adding noise with standard deviation 1/200th of the mean value)
increased mean absolute error of coalescence time inference from around 0.05 to 0.5.
Again, allowing multiple coalescence rates
or using resistance distance inference were highly inaccurate in all cases,
with mean absolute errors of above 0.5.

@@

We then test the method's ability to locate barriers to migration
on the $5 \times 3$ grid with asymmetric gene flow shown in Figure \ref{fig:5x3b_grid}. 
Each nonzero movement rate is determined randomly as before,
and coalescence rates are set to 1 everywhere.
Figure \ref{fig:5x3b_post_coalvcom} shows posterior distributions of the movement rates
for both coalescence time and resistance distance inference
on this graph, with noise standard deviation equal to
1/1000 of the mean coalescence time (left)
and 1/100 of the mean coalescence time (right).
Although increased noise in the data increased uncertainty in the estimated values,
coalescence time inference correctly inferred not only the location of the barrier
but also the migration rates along all other edges within a reasonable margin of error.
Resistance distance inference also identified the barrier,
but was much less accurate for the other values.
We also explored making the problem harder,
by first (a) dropping data corresponding to populations 2, 5, 7, and 15;
and then (b) allowing a separate coalescence rate for each location,
in both cases using a noise level of 1/100.
The resulting posterior distributions are shown in Figure \ref{fig:5x3boxplots_mult_gam},
and show substantially increased error,
but only in movement rates to the removed locations.
As before, uncertainty was increased when multiple coalescence rates were allowed,
but not as strongly as before, likely because the problem is less close to being underconstrained,
having 105 equations and 59 unknowns.
